The Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams

### 2016-06-28 oblique.tex
updated intro

The \Fermi satellite and ground based imaging atmospheric Cerenkov telescopes such as
% Phil: I added "imaging atmospheric" to distinguish this class of
% telescopes from imaging water CTs, e.g., MILAGRO; I am not aware of
% any space based IACT apart from a planned telescope on the ISS, but
% who knows whether this will ever fly
H.E.S.S., MAGIC, and VERITAS\footnote{High Energy
  Stereoscopic System, Major Atmospheric Gamma Imaging Cerenkov Telescope, Very
  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the ultra-high
energy Universe is teeming with energetic very high-energy gamma-ray 
(VHEGR, $E > 100\,\GeV$) sources, the extragalactic component of which mainly consists of TeV
blazars with a minority population of other sources  such as radio galaxies
and starburst galaxies.  These VHEGR observations have proven to be useful in constraining mechanisms 
of particle acceleration
\citep[see, e.g., ][]{Pagl_etal:96,Domi-Torr:05,Thom-Quat-Waxm:07,Pers-Reph-Arie:08,deCe-Torr-Rodr:09,Reph-Arie-Pers:10,Lack_etal:10} and
black hole jets
\citep[see, e.g., ][]{Jones+1974,Ghisellini+1989,Ghisellini+2008,Tavecchio+2008,Ghisellini+2009}.
% Phil: I cut this part because it is written as a fact (while it is not) and the same long author list is repeated in the next paragraph (the \qc command did not work with the \citep command):
% and intergalactic magnetic fields (IGMF)
% \citep[IGMF; ][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.

It has recently been argued by a number authors
that the {\it non-observation} of $\sim 1-100\,\GeV$ flux by the \Fermi satellite toward these TeV blazars provides lower bounds on the strength of the  intergalactic magnetic field (IGMF) of $B\gtrsim 10^{-19}-10^{-15}\,\G$ \citep{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.
To understand how these claimed limits arise, we begin by revisiting the physics of VHEGR propagation in the universe.

As these VHEGR photon propagate through the universe, they pair produce off the extragalactic background light (EBL) and produce ultrarelativistic electron-positron beams  with typical Lorentz factors of $10^5$--$10^7$.  
Within the current epoch, the mean free path for these photons is $\Dpp\sim 800 \,E_{\TeV}^{-1}\,\Mpc$, where $E_{\TeV} = E/1\,\TeV$ is the energy of the photon in \TeV.  At high redshifts, $z\geq 1$, this drops to $D_{\rmn{pp}}\sim 35\, E_{\TeV}^{-1}\,\Mpc$ due to the larger number density of EBL photons.  Because these typical mean free paths are much shorter than the Hubble length, the universe is optically thick to VHEGRs. This effect is seen in the relative nearness of TeV blazars ($z\sim 0.1$) and the observed softening of their VHE spectra with increasing distance.

Previously, it has been  assumed that the resultant electron-positron pairs lose energy through inverse-Compton scattering the cosmic microwave background (CMB) and EBL. This yields an inverse-Compton cascade (ICC) that reprocesses TeV emission to gamma-rays with energies $\lesssim 100 \GeV$, for which the mean free path is comparable to the size of the universe.    
Hence, associated with every VHEGR source should be a $\sim 100\,\GeV$ source.  It is the {\it non-observation} of this $100\,\GeV$ flux, presumably due to the magnetic deflection of the pairs, that is used to constraint the strength of a putative IGMF.

@@

Ultimately, these conclusions rest upon the assumption that ICCs dominate the evolution of the ultra-relativistic pairs.  
However, we have strongly challenged this assumption in a recent series of papers, \citet[][hereafter paper I, II, and III respectively]{BCP,CBP, PCB}.  In particular, we find that ultrarelativistic pair beam that results from TeV photon annihilation on the EBL is subject to virulent plasma beam instabilities that sap the energy of the pair beam and convert it to heat in the intergalactic medium (IGM) before ICCs have a chance to act.

In Paper I, we showed that despite the extraordinarily dilute nature of this ultra-relativistic pair plasma, a variety of plasma instabilities grow on timescales short in comparison to the inverse-Compton cooling time. Of the three instabilities that we studied --- the beam-plasma or two-stream instability\footnote{In order to follow the plasma literature more closely, henceforth we will refer to what we have called the ``two-stream'' instability as the ``beam-plasma'' instability}, the Weibel instability, and the ``oblique'' instability  ---  the most
important of which is the ``oblique'' instability \citep{Bret-Firp-Deut:04}
Because the pairs lose their energy primarily via depositing it as heat into the IGM, the ICCs critical to the above arguments commonly employed to provide constraints upon the IGMF and high-redshift blazar population are necessarily precluded.  In the latter case in particular, the \Fermi limits on the EGRB and high-energy blazar statistics allow a large and evolving blazar population more similar to that found for other AGN.

Assuming that these plasma-instability mechanisms drive ultra-relativistic pairs to deposit their energy as heat in
the IGM, we explored the impact of this heating on the
thermodynamics of the IGM in Paper II and the effect on this additional heating on structure formation in the universe in Paper III.  
By correcting for the selection effects of the current pencil beam VHEGR observations using
the all-sky monitoring of the \Fermi satellite, we argued that 
luminosity density in VHEGRs is of order $0.2\%$ of the quasar
luminosity density in Paper II.  Because of the greater conversion efficiency of TeV photon energy to heat (as compared to photoheating), we showed that this heating dominates the
photoheating rate at low $z$. This steady volumetric homogeneous heating naturally produces an inverted temperature-density relation and, with only a minor rescaling of the empirically determined TeV blazar
heating normalization, reproduces the inferred
inverted temperature-density relation at $z=2-3$ \citep{Bolton+09,Viel+09}, a feature that standard reionization models have proven unable to reproduce \citep{McQuinn+09,Bolton+09}.

Following the work of Paper II, we studied the effects of blazar heating in a
more detailed hydrodynamic model of structure formation 
\citep{Puchwein+2011}. Using the same blazar heating prescription as in Paper
II, we showed that the optical depth of the \Lya\ forest is reproduced using a
H{\sc i} photoionization rate of $\Gamma_{\rm H\mathsc{i}} \approx 5\times
10^{-13}\,{\rm s}^{-1}$ or equivalently using the inferred ionizing background
from \citet{FG+09}.  We also demonstrated that the low-density IGM again
possesses an inverted temperature-density relation, confirming the results of
Paper II in greater detail.  In addition, we showed that a blazar heated
universe not only matches the one- and two-point statistics of the
  high redshift Ly$\alpha$ forest but also the linewidth and column depth
  distribution. This excellent quantitative agreement is achieved
without tuning the photoionizing background and is due specifically to the
excess energy injection into the low-density IGM from TeV blazars.

Finally in Paper III, we explored the result of this additional IGM heating upon
the formation of structure in the Universe.  In particular, we showed that
the injection of entropy into the IGM by TeV blazars contributes to
developing a redshift dependent entropy floor for galaxy clusters and groups at
$z\lesssim 2$ and suppresses the formation of dwarfs.  The redshift dependent nature of TeV blazar heating suggests a large injection of entropy around $z\sim 1$, boosting the entropy of late forming groups and clusters.  This predicted enhanced entropy
of young groups is consistent with recent observations 
that show optically bright, and therefore young, groups and
clusters are X-ray dim ---  that is having a lower gas density due to a
raised entropy floor.  We also found that TeV
blazar heating suppresses the formation of late forming dwarfs both in
galactic halos, which is potentially important for the missing satellite problem \citep{Krav:10}, and
in voids, which is likewise potentially important for the void phenomenon \citep{Peebles2001}, by raising the
temperature of the IGM such that gas cannot collapse to form galaxies.

These potential implications of blazar heating rely on the assumption that the nonlinear state of these plasma instabilities efficiently convert the kinetic energy of the beam and convert it into heat.  In our previous work, we appealed to nonlinear simulations conducted in a radically different regime to argue for such a conversion.  However, a full nonlinear study is called for.  We begin this task in this initial work where we study the distribution function of the electron-positron pairs that are produced from VHEGR-EBL photon annihilation.  We study the evolution of a distribution function under Lorentz transformations to develop an analytic understanding of how the perpendicular and parallel velocity dispersions transform under boosts.  Using this understanding, we then develop a simple description of the distribution function of the beam, which we then use to calculate the unstable modes analytically in both the reactive (hydrodynamic) and kinetic regimes. Here the reactive instability refers to the instability where the entire beam participates in the instability.  In particular, all the beam particles are resonant with the unstable wave on a timescale longer than the growth time of the instability.  The reactive instability is also referred to as the hydrodynamic instability since the instability can be derived from the fluid equations instead of kinetic theory.  On the other hand, in the kinetic regime, only a fraction of the beam particles are resonant with the beam over a the growth time of the instability, which reduces the growth rate compared to the reactive instability for the same beam density and beam Lorentz factor.  We recover the well-known results for the reactive regime for both the beam-plasma and ``oblique'' modes.  We also derive the growth rate for these two instabilities in the kinetic regime and delineate the range of applicability for both the reactive and kinetic cases and apply these results for ultrarelativistic \epm\ pair beams.

@@

In the COM frame of the beam, we assume that the distribution function is isotropic, such that $f=f(E)$ is just a function of energy.  We model this distribution as a relativistic thermal distribution:
\begin{equation}\label{eq:relativistic maxwellian}
f\propto \exp\left(-\frac E T\right),
\end{equation}
 where $E$ and $T$ are the dimensionless energy and temperature in terms of a particles rest mass.  In the non-relativistic case, this reduces to the Maxwell-Boltzmann distribution, while the relativistic version is known as the Maxwell-Juttner distribution \citep{1911AnP...340..145J}.

@@

The Vlasov equation for each species is
\begin{equation}\label{eq:linearize vlasov}
  \frac {\partial f_s}{\partial t} + \vel_s \cdot \grad f_s + 
  q_s \left({\Efield} + \frac {\vel_s} {c} \times {\Bfield} \right)
 \cdot\grad_p f_s = 0,
\end{equation}
where $\vel_s=\pmom_s/\gamma_s m_e$ and  $\gamma_s = 1/\sqrt{1-v_s^2/c^2} $
Here, $s$ is the species label, $+$ for positrons and $-$ for electrons,
with $q_{\pm} = \pm e$.  %\ac{[$v$ vs. $\vel$ vs. $\beta$ problems above.]}
Upon linearizing this in small perturbations about a background distribution, i.e., setting 
$f_s\rightarrow f_{0s} + \delta f_s$, $\Bfield \rightarrow \delta \Bfield$ and $\Efield \rightarrow \delta \Efield$,  we obtain, 
\begin{equation}\label{eq:vlasov}
  \frac {\partial\delta f_s}{\partial t} + {\vel_s} \cdot \grad\delta f_s  
  +  {q_s} \left(\delta \Efield + \frac {\vel_s} {c} \times \delta \Bfield \right)\cdot\grad_p f_{0s} = 0.
 \end{equation}
The plasma couples to the field through the Maxwell equations
\begin{eqnarray}\label{eq:maxwell}
\grad \times \delta \Efield &=& -\frac 1 c \frac {\partial \delta \Bfield} {\partial t},\\
\grad \times \delta \Bfield &=& \frac {4\pi} {c} \delta\jvec + \frac{1}{ c} \frac {\partial \delta \Efield} {\partial t},\label{eq:maxwell2}
\end{eqnarray}
where $\delta\jvec = \sum_s q_s\int \vel_s \delta f_s d^3 p$ is the linear current density perturbation.

@@

For $k_x = 0 \rightarrow Z_x = 0$, this reduces to the beam-plasma growth rate, which is
\begin{equation}\label{eq:growth rate bp}
 \Gamma = \Gamma_{\rm TS} \equiv \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\frac{\omega_{p,t}}{\gamma},
\end{equation}
which we denote the beam-plasma or ``two-stream'' growth rate.
For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by Bret et al.  Indeed for $\gamma \gg 1$ and $Z_x \sim 1$, the growth rate approaches oblique growth rate:
\begin{equation}
 \Gamma = \Gamma_{\rm ob} \equiv \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\frac{\omega_{p,t}}{\gamma^{1/3}},
\end{equation}
which is much faster than the beam-plasma growth rate, $\Gamma_{\rm TS}$

@@

Equation (\ref{eq:dispersion relation reactive}) can also be solved numerically in terms of $k_x$ and $k_z$.  Here let us specialize to the case of $k_x = 0$, i.e., the beam-plasma case.  In this case, we have 
\begin{equation}\label{eq:dispersion relation beam-plasma}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2} = 0,
\end{equation}
which we can numerically solve in terms of $\omega/\omega_{p,t}$, $k_z\lambda_D$, $v_b/c$, and $n_b/n_t$, where $\lambda_D = c/\omega_{p,t}$ is the skin depth.  In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for representative case of $v_b/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.    For $k_z\lambda_D = 1$, the growth rate reaches it maximum of $\Gamma_{\rm max}$ and the real part of the frequency is $\Re(\omega) = \omega_{p,t}$, which is plasma oscillation frequency.  This wave would exist in the absence of a tenuous beam.  However, as we move away from this frequency toward lower $k_z$, we still find substantial growth, with $\Gamma \approx 0.1 \Gamma_{\rm max}$ as $k_z \approx 0.9\lambda_D^{-1}$.  Interestingly, the real part of the unstable wave has a phase velocity, $v_{\rm ph} = \Re(\omega)/k = c$, which is still in resonance with the beam.  This unstable wave does not exist in the absence of the destabilizing beam and does not correspond to any wave in the unperturbed system.

@@

To begin let us return to the reactive instability.  For the growth rate of the reactive instability (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion $\rightarrow 0$.   To reach this limit, we demand
\begin{equation}
 \kvec\cdot\Delta \vel \ll \omega - \kvec\cdot\vel_b.
\end{equation}
Taking $\kvec$ to be $\omega_{p,t}/\vel_b$, so that the RHS is $\approx \Gamma$, we find
\begin{equation}\label{eq:oblique reactive regime}
 \frac{\Delta v_{\perp}}{v_b} \ll \left(\frac{n_b}{\gamma_b n_t}\right)^{1/3}
%\left(1 - \beta_b^2\cos^2\Psi\right)^{1/3}, 
\end{equation}
where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$.  Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate (\ref{eq:growth rate reactive}).  For $Z_x \ll \gamma^{-2}$, we recover the condition for the relativistic, reactive beam-plasma instability \citep{Boyd}
\begin{equation}\label{eq:bp reactive regime}
\frac{\Delta v_{\parallel}}{v_b} \ll \gamma_b^{-1}\left(\frac{n_b}{n_t}\right)^{1/3}.
\end{equation}

>>>>

This excess heating of the IGM may resolve a variety
of cosmological puzzles, including naturally explaining
  anomalies in the statistics of the high-redshift Ly$\alpha$ forest
\citep{paperIV} and potentially explaining a number of the X-ray
  properties of groups and clusters and anomalies in galaxy formation on
  the scale of dwarfs \citep{paperIII,Lu+2013}.  We
  have recently shown that \textit{if} the IC halos are ignored, it is possible
  to quantitatively reproduce the redshift and flux distributions
  of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
  spectrum above 3~GeV simultaneously with a unified model of AGN evolution
  \citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
  circumstantial evidence for the presence of virulent plasma beam
  instabilities.

The \Fermi satellite and ground based imaging atmospheric Cerenkov telescopes such as
% Phil: I added "imaging atmospheric" to distinguish this class of
% telescopes from imaging water CTs, e.g., MILAGRO; I am not aware of
% any space based IACT apart from a planned telescope on the ISS, but
% who knows whether this will ever fly
H.E.S.S., MAGIC, and VERITAS\footnote{High Energy
  Stereoscopic System, Major Atmospheric Gamma Imaging Cerenkov Telescope, Very
  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the ultra-high
energy Universe is teeming with energetic very high-energy gamma-ray 
(VHEGR, $E > 100\,\GeV$) sources, the extragalactic component of which mainly consists of TeV
blazars with a minority population of other sources  such as radio galaxies
and starburst galaxies. These extragalactic VHEGR emitters produce TeV photons that are greatly attenuated via annihilation upon soft photons in the extragalactic background light
(EBL) and produce pairs \citep[see, e.g.,][]{Goul-Schr:67,Sala-Stec:98,Nero-Semi:09}.

It has been assumed that these ultrarelativistic pairs produced by VHEGR annihilation 
lose energy exclusively through inverse-Compton (IC) scattering off of the cosmic microwave
background (CMB),   transferring the energy of the original VHEGR to gamma-rays with
energies $\lesssim100\,\GeV$.  The absence of observed secondary IC emission leads a number of
authors to argue that this \textit{lack} places lower bounds upon the
intergalactic magnetic field
\citep[IGMF; see, e.g.,][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Taka_etal:11,Dola_etal:11} with 
typical numbers ranging from $10^{-19}\,\G$ to $10^{-15}\,\G$.

@@

These two important conclusions depends on IC cooling dominating the evolution of the ultra-relativistic pairs.  However, it was recently found that plasma instabilities driven by the
ultrarelativistic pair beams  likely is the dominant cooling mechanism
\citep{paperI,Schlickeiser+12,Schlickeiser+13}, depositing this energy
as heat in the intergalactic medium \citep{paperII,paperIII}.  Therefore, the lack of an observed
 IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
\citep{paperI,Schlickeiser+12,Schlickeiser+13}.

% Phil: I cut this part because it is written as a fact (while it is not) and the same long author list is repeated in the next paragraph (the \qc command did not work with the \citep command):
% and intergalactic magnetic fields (IGMF)
% \citep[IGMF; ][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.

The deposition of the kinetic energy into the IGM via plasma instabilities produces excess heating, which over cosmological time, may resolve a variety
of puzzles, including explaining
anomalies in the statistics of the high-redshift Ly$\alpha$ forest
\citep{paperIV} and potentially explaining a number of the X-ray
properties of groups and clusters and anomalies in galaxy formation on
the scale of dwarfs \citep{paperIII,Lu+2013}.  We
have recently shown that \textit{if} the IC halos are ignored, it is possible
to quantitatively reproduce the redshift and flux distributions
of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
spectrum above 3~GeV simultaneously with a unified model of AGN evolution
\citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
circumstantial evidence for the presence of virulent plasma beam
instabilities.

These potential implications of blazar heating rely on an understanding of the linear and nonlinear physics of these plasma instabilities.  Recent work in this area has been inconclusive.  For instance, \citet{Miniati+12} argued that these instabilities are physically irrelevant for the cooling of these pair beams because they would saturate at a very low level due to nonlinear Landau damping (NLD).  However, \citet{Chang+14} showed that a detailed calculation of NLD showed that these plasma processes remain dominant.  In addition, \citet{Siron+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.  It is unclear, however, if the conclusions of their work is robust as it is unclear if these numerical setup is adequate (e.g. Mohamad's paper).

However, a full nonlinear study is called for.  We begin this task in this initial work where we study the distribution function of the electron-positron pairs that are produced from VHEGR-EBL photon annihilation.  We study the evolution of a distribution function under Lorentz transformations to develop an analytic understanding of how the perpendicular and parallel velocity dispersions transform under boosts.  Using this understanding, we then develop a simple description of the distribution function of the beam, which we then use to calculate the unstable modes analytically in both the reactive (hydrodynamic) and kinetic regimes. Here the reactive instability refers to the instability where the entire beam participates in the instability.  In particular, all the beam particles are resonant with the unstable wave on a timescale longer than the growth time of the instability.  The reactive instability is also referred to as the hydrodynamic instability since the instability can be derived from the fluid equations instead of kinetic theory.  On the other hand, in the kinetic regime, only a fraction of the beam particles are resonant with the beam over a the growth time of the instability, which reduces the growth rate compared to the reactive instability for the same beam density and beam Lorentz factor.  We recover the well-known results for the reactive regime for both the beam-plasma and ``oblique'' modes.  We also derive the growth rate for these two instabilities in the kinetic regime and delineate the range of applicability for both the reactive and kinetic cases and apply these results for ultrarelativistic \epm\ pair beams.

@@

In the COM frame of the beam, we assume that the distribution function is isotropic, such that $f=f(E)$ is just a function of energy.   This equilibrium energy distribution of a relativistic thermal plasma gas is
\begin{equation}\label{eq:relativistic maxwellian}
f\propto \exp\left(-\frac E T\right),
\end{equation}
 where $E$ and $T$ are the dimensionless energy and temperature in terms of a particles rest mass.  In the non-relativistic case, this reduces to the Maxwell-Boltzmann distribution, while the relativistic version is known as the Maxwell-Juttner distribution \citep{1911AnP...340..145J}.

@@

The Vlasov equation for each species is
\begin{equation}\label{eq:linearize vlasov}
  \frac {\partial f_s}{\partial t} + \vel_s \cdot \grad f_s + 
  q_s \left({\Efield} + \frac {\vel_s} {c} \times {\Bfield} \right)
 \cdot\grad_p f_s = 0,
\end{equation}
where $\vel_s=\pmom_s/\gamma_s m_e$ and  $\gamma_s = 1/\sqrt{1-v_s^2/c^2} $
Here, $s$ is the species label, $+$ for positrons and $-$ for electrons,
with $q_{\pm} = \pm e$.  %\ac{[$v$ vs. $\vel$ vs. $\beta$ problems above.]}
Upon linearizing this in small perturbations about a background distribution, i.e., setting 
$f_s\rightarrow f_{0s} + \delta f_s$, $\Bfield \rightarrow \delta \Bfield$ and $\Efield \rightarrow \delta \Efield$,  we obtain, 
\begin{equation}\label{eq:vlasov}
  \frac {\partial\delta f_s}{\partial t} + {\vel_s} \cdot \grad\delta f_s  
  +  {q_s} \left(\delta \Efield + \frac {\vel_s} {c} \times \delta \Bfield \right)\cdot\grad_p f_{0s} = 0.
 \end{equation}
The plasma couples to the field through the Maxwell equations
\begin{eqnarray}\label{eq:maxwell}
\grad \times \delta \Efield &=& -\frac 1 c \frac {\partial \delta \Bfield} {\partial t},\\
\grad \times \delta \Bfield &=& \frac {4\pi} {c} \delta\jvec + \frac{1}{ c} \frac {\partial \delta \Efield} {\partial t},\label{eq:maxwell2}
\end{eqnarray}
where $\delta\jvec = \sum_s q_s\int \vel \delta f_s d^3 p$ is the linear current density perturbation.

@@

For $k_x = 0 \rightarrow Z_x = 0$, this reduces to the beam-plasma growth rate, which is
\begin{equation}\label{eq:growth rate bp}
 \Gamma = \Gamma_{\rm TS} \equiv \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\frac{\omega_{p,t}}{\gamma},
\end{equation}
which we denote the beam-plasma or ``two-stream'' growth rate.
For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by Bret et al.  Indeed for $\gamma \gg 1$ and $Z_x \rightarrow \infty$, the growth rate approaches oblique growth rate:
\begin{equation}
 \Gamma = \Gamma_{\rm ob} \equiv \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\frac{\omega_{p,t}}{\gamma^{1/3}},
\end{equation}
which is much faster than the beam-plasma growth rate, $\Gamma_{\rm TS}$

@@

Equation (\ref{eq:dispersion relation reactive}) can also be solved numerically in terms of $k_x$ and $k_z$.  Here let us specialize to the case of $k_x = 0$, i.e., the beam-plasma case.  In this case, we have 
\begin{equation}\label{eq:dispersion relation beam-plasma}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2} = 0,
\end{equation}
which we can numerically solve in terms of $\omega/\omega_{p,t}$, $k_z\lambda_D$, $v_b/c$, and $n_b/n_t$, where $\lambda_D = c/\omega_{p,t}$ is the skin depth.  In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for representative case of $v_b/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.    For $k_z\lambda_D = 1$, the growth rate reaches it maximum of $\Gamma_{\rm max}$ and the real part of the frequency is $\Re(\omega) = \omega_{p,t}$, which is plasma oscillation frequency.  This wave would exist in the absence of a tenuous beam.  However, as we move away from this frequency toward lower $k_z$, we still find substantial growth, with $\Gamma \approx 0.1 \Gamma_{\rm max}$ as $k_z \approx 0.9\lambda_D^{-1}$.  Interestingly, the real part of the unstable wave has a phase velocity, $v_{\rm ph} = \Re(\omega)/k = c$, which is still in resonance with the beam.  This unstable wave does not exist in the absence of the destabilizing beam and does not correspond to any wave in the unperturbed target plasma.

@@

To begin let us return to the reactive instability.  For the growth rate of the reactive instability (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion $\rightarrow 0$.   To reach this limit, we demand
\begin{equation}
 \kvec\cdot\Delta \vel \ll \omega - \kvec\cdot\vel_b.
\end{equation}
Taking $\kvec$ to be $\omega_{p,t}/\vel_b$, so that the RHS is $\approx \Gamma$, we find
\begin{equation}\label{eq:oblique reactive regime}
 \frac{\Delta v_{\perp}}{v_b} \ll \left(\frac{n_b}{\gamma_b n_t}\right)^{1/3}
%\left(1 - \beta_b^2\cos^2\Psi\right)^{1/3}, 
\end{equation}
where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$.  Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate (\ref{eq:growth rate reactive}).  For $Z_x \ll \gamma^{-2}$, we recover the condition for the relativistic, reactive beam-plasma instability:
\begin{equation}\label{eq:bp reactive regime}
\frac{\Delta v_{\parallel}}{v_b} \ll \gamma_b^{-1}\left(\frac{n_b}{n_t}\right)^{1/3}.
\end{equation}
